---
title: "Campus Culture"
---

<!-- Setup -->

```{r}
source('setup.R')
```

### Perceptions of Open Source Culture On Campus

**`r round(length(survey_results$QID24[survey_results$QID24 == 'Very valuable']) / nrow(survey_results), 2) * 100`%** of respondents said that having a vibrant open source culture is **"very valuable"**.

```{r}
c1_df <- survey_results |> 
  mutate(QID4 = ifelse(is.na(QID4), 'Unafilliated', QID4),
         QID24 = factor(QID24, levels = c(
          'Very valuable',
          'Some value',
          'Neutral',
          'No value'
         ))) |> 
  rename(`Respondent Type` = QID4) |> 
  group_by(QID24, `Respondent Type`) |> 
  summarise(Count = n(),
            Percent = Count / nrow(survey_results)) 

c2_df <- survey_results |> 
  mutate(QID4 = ifelse(is.na(QID4), 'Unafilliated', QID4),
         QID23 = factor(QID23, levels = c(
          'Strongly agree',
          'Somewhat agree',
          'Neither agree nor disagree',
          'Somewhat disagree',
          'Strongly disagree'
         ))) |> 
  rename(`Respondent Type` = QID4) |> 
  group_by(QID23, `Respondent Type`) |> 
  summarise(Count = n(),
            Percent = Count / nrow(survey_results)) 
```

```{r}
c1_df |>
  plot_ly(
    x = ~ QID24,
    y = ~ Percent,
    color = ~ `Respondent Type`,
    colors = viridis_pal(option = "D")(length(c1_df$`Respondent Type`))
  ) |>
  add_bars() |>
  layout(
    barmode = 'stack',
    plot_bgcolor = background_color,
    paper_bgcolor = background_color,
    title = paste0('Do you see value in having a vibrant culture around open source software?'),
    xaxis = list(title =  '', tickangle = 25),
    yaxis = list(
      zerolinecolor = '#ffff',
      zerolinewidth = 2,
      gridcolor = 'ffff',
      tickformat = ".1%"
    )
  )
```

In comparison, only **`r round(length(survey_results$QID23[survey_results$QID23 %in% c('Strongly agree', 'Somewhat agree')]) / nrow(survey_results), 2) * 100`%** agreed that there is a vibrant open source culture at `r uni_name`.

```{r}
c2_df |>
  plot_ly(
    x = ~ QID23,
    y = ~ Percent,
    color = ~ `Respondent Type`,
    colors = viridis_pal(option = "D")(length(c1_df$`Respondent Type`))
  ) |>
  add_bars() |>
  layout(
    barmode = 'stack',
    plot_bgcolor = background_color,
    paper_bgcolor = background_color,
    title = paste0('"', uni_name, ' has a vibrant open source culture"'),
    xaxis = list(title =  ''),
    yaxis = list(
      zerolinecolor = '#ffff',
      zerolinewidth = 2,
      gridcolor = 'ffff',
      tickformat = ".1%"
    )
  )
```

**`r round(length(survey_results$QID19[survey_results$QID19 == 'Yes']) / nrow(survey_results), 2) * 100`%** of respondents agreed that **"it makes sense for the university to contribute to open source software that is vital to its educational and research enterprise"** (`r round(length(survey_results$QID19[survey_results$QID19 == 'Not sure']) / nrow(survey_results), 2) * 100`% were "unsure," and `r round(length(survey_results$QID19[survey_results$QID19 == 'No']) / nrow(survey_results), 2) * 100`% disagreed).

### Open Source Training On Campus

**`r round(length(survey_results$QID25[survey_results$QID25 == 'Yes']) / nrow(survey_results), 2) * 100`%** of respondents said that they have received at least some formal training or education on open source software during their time at `r uni_name`. These respondents were distributed by role and affiliation as below:

::: {.panel-tabset}

```{r}
c3_df <- survey_results |> 
  mutate(QID4 = ifelse(is.na(QID4), 'Unafilliated', QID4)) |> 
  rename(`Respondent Type` = QID4) |> 
  group_by(QID25, `Respondent Type`) |> 
  summarise(Count = n(),
            Percent = Count / nrow(survey_results)) 

c4_df <- survey_results |> 
  mutate(QID8 = ifelse(is.na(QID8), 'Unafilliated', QID8)) |> 
  rename(`Affiliation` = QID8) |> 
  group_by(QID25, `Affiliation`) |> 
  summarise(Count = n(),
            Percent = Count / nrow(survey_results)) 
```

### Role

```{r}
c3_df |>
  plot_ly(
    x = ~ QID25,
    y = ~ Percent,
    color = ~ `Respondent Type`,
    colors = viridis_pal(option = "D")(length(c3_df$`Respondent Type`))
  ) |>
  add_bars() |>
  layout(
    barmode = 'stack',
    plot_bgcolor = background_color,
    paper_bgcolor = background_color,
    xaxis = list(title = paste0('Have received open source training at ', uni_name)),
    yaxis = list(
      zerolinecolor = '#ffff',
      zerolinewidth = 2,
      gridcolor = 'ffff',
      tickformat = ".1%"
    )
  )
```

### Affiliation

```{r}
c4_df |>
  plot_ly(
    x = ~ QID25,
    y = ~ Percent,
    color = ~ `Affiliation`,
    colors = viridis_pal(option = "D")(length(c4_df$`Affiliation`))
  ) |>
  add_bars() |>
  layout(
    barmode = 'stack',
    plot_bgcolor = background_color,
    paper_bgcolor = background_color,
    xaxis = list(title = paste0('Have received open source training at ', uni_name)),
    yaxis = list(
      zerolinecolor = '#ffff',
      zerolinewidth = 2,
      gridcolor = 'ffff',
      tickformat = ".1%"
    ),
    hoverlabel = list(namelength = -1)
  )
```

:::

**`r round(length(survey_results$QID28[survey_results$QID28 == 'Yes']) / nrow(survey_results), 2) * 100`%** of respondents said that they "would like to see more training, education, or support for learning how to contribute to open source project" and **`r round(length(survey_results$QID29[survey_results$QID29 == 'Yes']) / nrow(survey_results), 2) * 100`%** of respondents expressed interest in potential open source training sessions and workshops organized by the Open Source Program Office.
