---
title: "Sentiments Around Open-Source Tools"
---

<!-- Setup -->

```{r}
source('setup.R')
```

### Value of Open Source Projects

When asked to characterize the importance of open source tools to their education, teaching, and research, respondents said the following:

```{r}
s1_df <- survey_results |> 
  mutate(QID4 = ifelse(is.na(QID4), 'Unafilliated', QID4),
         QID16 = factor(QID16,
                        levels = c(
                          'Critical/Vital/Essential',
                          'Valuable',
                          'Neutral',
                          'Little importance',
                          'Irrelevant'
                        ))) |> 
  rename(`Respondent Type` = QID4) |> 
  group_by(QID16, `Respondent Type`) |> 
  summarise(Count = n(),
            Percent = Count / nrow(survey_results))

s1_df |>
  plot_ly(
    x = ~ QID16,
    y = ~ Percent,
    color = ~ `Respondent Type`,
    colors = viridis_pal(option = "D")(length(s1_df$`Respondent Type`))
  ) |>
  add_bars() |>
  layout(
    barmode = 'stack',
    plot_bgcolor = background_color,
    paper_bgcolor = background_color,
    xaxis = list(title = ''),
    yaxis = list(
      zerolinecolor = '#ffff',
      zerolinewidth = 2,
      gridcolor = 'ffff',
      tickformat = ".1%"
    )
  )
```

Survey takers were also asked how much they agree that 1) open source tools are a valuable form of research output and 2) open source tools are an important way to translate research into entrepreneurship and innovation.

```{r}
s2_df <- survey_results |> 
  mutate(QID4 = ifelse(is.na(QID4), 'Unafilliated', QID4),
         QID20 = factor(QID20, levels = c(
          'Strongly agree',
          'Somewhat agree',
          'Neither agree nor disagree',
          'Somewhat disagree',
          'Strongly disagree'
         ))) |> 
  rename(`Respondent Type` = QID4) |> 
  group_by(QID20, `Respondent Type`) |> 
  summarise(Count = n(),
            Percent = Count / nrow(survey_results)) 

s3_df <- survey_results |> 
  mutate(QID4 = ifelse(is.na(QID4), 'Unafilliated', QID4),
         QID21 = factor(QID21, levels = c(
          'Strongly agree',
          'Somewhat agree',
          'Neither agree nor disagree',
          'Somewhat disagree',
          'Strongly disagree'
         ))) |> 
  rename(`Respondent Type` = QID4) |> 
  group_by(QID21, `Respondent Type`) |> 
  summarise(Count = n(),
            Percent = Count / nrow(survey_results)) 
```

::: {.panel-tabset}

### "OS is a valuable research output"

```{r}
s2_df |>
  plot_ly(
    x = ~ reorder(QID20, desc(Percent)),
    y = ~ Percent,
    color = ~ `Respondent Type`,
    colors = viridis_pal(option = "D")(length(s2_df$`Respondent Type`))
  ) |>
  add_bars() |>
  layout(
    barmode = 'stack',
    plot_bgcolor = background_color,
    paper_bgcolor = background_color,
    xaxis = list(title = ''),
    yaxis = list(
      zerolinecolor = '#ffff',
      zerolinewidth = 2,
      gridcolor = 'ffff',
      tickformat = ".1%"
    )
  )
```

### "OS translates research into innovation"

```{r}
s3_df |>
  plot_ly(
    x = ~ reorder(QID21, desc(Percent)),
    y = ~ Percent,
    color = ~ `Respondent Type`,
    colors = viridis_pal(option = "D")(length(s3_df$`Respondent Type`))
  ) |>
  add_bars() |>
  layout(
    barmode = 'stack',
    plot_bgcolor = background_color,
    paper_bgcolor = background_color,
    xaxis = list(title = ''),
    yaxis = list(
      zerolinecolor = '#ffff',
      zerolinewidth = 2,
      gridcolor = 'ffff',
      tickformat = ".1%"
    )
  )
```

:::

### Benefits of Open Source Tools in Academic Settings

We asked survey respondents to choose among a list of potential benefits of using open-source tools in their work. Of the survey-takers who identified benefits, the percentage of respondents who identified each benefit are shown below:

```{r}

s4 <- survey_results |> 
  select(ResponseId, QID18) |> 
  separate_rows(QID18, sep = ',') |> 
  group_by(Benefit = QID18) |> 
  summarise(count = n(),
            pct = count/nrow(survey_results)) |> 
  mutate(Benefit = ifelse(is.na(Benefit), 'None', Benefit)) |> 
  arrange(desc(pct)) |> 
  ggplot() +
    geom_segment(
      aes(x = reorder(Benefit, pct), xend = reorder(Benefit, pct), y = 0, yend = pct), color = primary_color
    ) +
    geom_point(
      aes(x = Benefit, y = pct, text = paste0(round(pct * 100, 2), '%')), 
      color = primary_color
    ) +
    theme_minimal() +
    scale_color_identity() +
    scale_y_continuous(labels = percent_format(), limits = c(0,1)) +
    coord_flip() +
    xlab("") +
    ylab("Percent of Respondents Agreeing with Benefits of Using Open Source Software") +
    lollipop_theme

ggplotly(s4, tooltip = c('text'))

```
